Manuscript Figures

### 2024-11-05 manuscript_figures.Rmd
Msystems resubmission

#Loading in mapping file
mapping = readxl::read_excel("../proteomics/sample_mapping.xlsx")

@@

#Loading in data
data = read_delim("../proteomics/fragpipe_output_fixed_labels/tmt-report/abundance_gene_MD.tsv")

@@

@@

#ggsave("02_figures/f3_infection_v_healthy_metabolomics/f3d.pdf",f3d,height = 8,width = 8,units = "in")
```

@@

@@

# ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4d.pdf",f4d,height = 8,width = 8)
```

@@

#ggsave("02_figures/f5_faecalis_v_faecium_metabolomics/f5b.pdf",f5b,height = 8,width = 8)
```

@@

```{r,fig.height = 4, fig.width = 4}
#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

@@

f6b = clusterProfiler::cnetplot(enrich, categorySize="pvalue",circular = FALSE,categorySize="pvalue",cex_label_category = 1,showCategory = 100,cex_label_gene = 0.5)+
  theme(legend.position = "none")

@@

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6b.pdf",f6b,height = 4, width = 4)

@@

f6c = clusterProfiler::cnetplot(enrich, categorySize="pvalue",
                              circular = FALSE,
                              categorySize="pvalue",
                              cex_label_category = 1,
                              showCategory = 100,
                              cex_label_gene = 0.5, )+
  theme(legend.position = "none")
f6c

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6c.pdf",f6c)
```

@@

clean_clinical = inner_join(pl,md_final, by = "tmt_id") %>% 
  # Don't need the healthy samples here
  filter(Condition != "Healthy") %>%
  #IGKC gets identified by EFS even though it is a terrible biomarker. Not sure why, just going to remove it
  filter(Gene != "IGKC") %>% 
   # - messes up EFS for some reason, very annoying. 
  mutate(Gene = gsub("-","_",Gene)) %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #Removing samples with NA.
  filter(!is.na(death_during_admission)) %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame() %>% 
  select(-sample_id)

@@

EFS
```

@@

top = make_violin_plot_metabolomics(data,ncol = 2,topn = topn[1:2],pname = "p.signif")
bottom = make_roc_metabolomics(data,ncol = 2)

f6f = ggpubr::ggarrange(top,bottom,nrow = 2)

@@

@@

## Plotting ###
s7c = stat_final %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~Gene,nrow = 2) +
  xlab("metadata_field")

@@

## Plotting ###
s7d = stat_final %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~compound_name_cleaned,nrow = 2) +
  xlab("metadata_field")

@@

```{r,fig.height = 10, fig.width = 30}
## Reading in the PTM data from GNPS ## 
ptm_data = read_delim("../metabolomics/ProteoSAFe-METABOLOMICS-SNETS-V2-3e73be4d-view_edges_delta_histogram/networkedges_selfloop/ae6d9d27f8684ad39b8e73b2d6f3da2e..selfloop",na = " ") %>% 
  mutate(DeltaMZ = abs(DeltaMZ))

@@

```{r}
arranged_annotations = dplyr::arrange(annotations,-n)

@@

```{r}
#total number of MS2 scans
ms2_scans = read_delim("../proteomics/gnps/clusterinfo/93fdf96f02e44389b503c53353712f6a.clusterinfo") %>% 
  group_by(`#Filename`) %>% 
  summarise(n = n())

27017 + 16665
```

>>>>

#Loading in mapping file
mapping = readxl::read_excel("00_r_input_data/proteomics_sample_mapping.xlsx")

@@

#Loading in data
data = read_delim("00_r_input_data/abundance_gene_MD.tsv")

@@

# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = Gene)

write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/healthy_prot_EFS_rank.csv")

@@

#ggsave("02_figures/f3_infection_v_healthy_metabolomics/f3d.pdf",f3d,height = 8,width = 8,units = "in")
# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = row_id)
write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/infected_metabolomics_EFS.csv")
```

@@

#test

pl %>% 
  select(BioReplicate,Condition) %>% 
  distinct() %>% 
  pull(Condition) %>% 
  table()

@@

# ggsave("02_figures/f4_faecalis_v_faecium_proteomics/f4d.pdf",f4d,height = 8,width = 8)
# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = Gene)

write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/EcB_prot_EFS_rank.csv")
```

@@

#ggsave("02_figures/f5_faecalis_v_faecium_metabolomics/f5b.pdf",f5b,height = 8,width = 8)

# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = row_id)
write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/EcB_metabolomics_EFS.csv")
```

@@

```{r,fig.height = 6, fig.width = 6}
#Getting all the go annotations 
go = read_csv("00_r_input_data/go_annotations.csv") %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs)) %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>% 
  select(-Entry) %>% 
  select(Entry = Gene,Gene.Ontology.IDs)

@@

f6b = clusterProfiler::cnetplot(enrich, categorySize="pvalue",circular = FALSE,categorySize="pvalue",cex_label_category = 0.8,showCategory = 100,cex_label_gene = 0.5)+
  theme(legend.position = "none")

@@

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6b.pdf",f6b,height = 6, width = 6)

@@

f6c = clusterProfiler::cnetplot(enrich, categorySize="pvalue",
                              circular = FALSE,
                              categorySize="pvalue",
                              cex_label_category = 0.8,
                              showCategory = 100,
                              cex_label_gene = 0.5, )+
  theme(legend.position = "none")
f6c

ggsave("02_figures/f6_prediction_of_clinical_outcome/f6c.pdf",f6c,width = 6, height = 6)
```

@@

clean_clinical = inner_join(pl,md_final, by = "tmt_id") %>% 
  # Don't need the healthy samples here
  filter(Condition != "Healthy") %>%
  #These get identified by EFS even though it is a terrible biomarker. Not sure why, just going to remove them
  filter(Gene != "IGKC") %>%  filter(Gene != "SOD1") %>%  filter(Gene != "CD5L") %>% filter(Gene != "IGHM") %>% 
   # - messes up EFS for some reason, very annoying. 
  mutate(Gene = gsub("-","_",Gene)) %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #Removing samples with NA.
  filter(!is.na(death_during_admission)) %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame() %>% 
  select(-sample_id)

@@

EFS

# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = Gene)

write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/mortality_prot_EFS_rank.csv")
```

@@

topvp = make_violin_plot_metabolomics(data,ncol = 2,topn = topn[1:2],pname = "p.signif")
bottom = make_roc_metabolomics(data,ncol = 2)

f6f = ggpubr::ggarrange(topvp,bottom,nrow = 2)

@@

# writing for shiny app
shiny_out = top %>% 
  tibble::rowid_to_column(var = "rank") %>% 
  dplyr::rename(Feature = row_id)

write_csv(shiny_out,"04_shiny_app/Biomarker_Explorer/www/mortality_metabolomics_EFS.csv")

@@

## Plotting ###
s7c = stat_final %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~factor(Gene,topn),nrow = 2) +
  xlab("metadata_field")

@@

  #ordering ids 
  ordered = read_csv("01_r_processed_data/top10_live_v_dead_metabolomics_biomarkers.csv") %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
    select(row_id,compound_name_cleaned) %>% 
    pull(compound_name_cleaned)

## Plotting ###
s7d = stat_final %>% 
  inner_join(gnps_annotations,by = "row_id") %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~factor(compound_name_cleaned,ordered),nrow = 2) +
  xlab("metadata_field")

@@

```{r,fig.height = 10, fig.width = 30, eval = FALSE}
## Reading in the PTM data from GNPS ## 
ptm_data = read_delim("../metabolomics/ProteoSAFe-METABOLOMICS-SNETS-V2-3e73be4d-view_edges_delta_histogram/networkedges_selfloop/ae6d9d27f8684ad39b8e73b2d6f3da2e..selfloop",na = " ") %>% 
  mutate(DeltaMZ = abs(DeltaMZ))

@@

```{r,eval = FALSE}
arranged_annotations = dplyr::arrange(annotations,-n)

@@

```{r,eval = FALSE }
#total number of MS2 scans
ms2_scans = read_delim("../proteomics/gnps/clusterinfo/93fdf96f02e44389b503c53353712f6a.clusterinfo") %>% 
  group_by(`#Filename`) %>% 
  summarise(n = n())
